### 2015-07-22 manuscript/BMC/bmc_article.tex
Updating defs / manuscript

The most well known tools for repeat identification ar RepeatMasker
and [Eddy's tool] [CITE], both of which work on the principle of a
library-based search.  That is, to final transposable elements of a
given family,

>>>>

The most well known tools for repeat identification ar RepeatMasker
and [Eddy's tool] [CITE], both of which work on the principle of a
library-based search.  That is, to find transposable elements of a
given family, the tools require some description of sequence in the
family (in the form of a model sequence for blasting, or a pHMM) that
is used as the basis for identification.  This begs the obvious
question: much like we ask how the snow plow driver gets to work, or
how dictinary authors look up the spelling of a word [CITE], we must ask how the
tools compile their libraries.  How do we compile the libraries of
transposible elements, or discover new transpoable element families.

Within mammalian speices we can, to some extent, rely on homology
relationships to port libraries across species.  A library description
of a human ALU sequence can likely be used for the initial annotation
of a newly sequence primate sequence, and many LINEs will extend to
rodent, or even more distantn mammalian species.  This does not hold
so well in plants.  In many cases TE composition of a given organism
is species-specific.  For example, a rice-based TE libarary will only identify 25\%
of the TEs in the maize genome. [CITE Jiang]

To solve this problem we turn to {\it de novo} TE identification
tools: tools identifying TEs though a genome using only the genome
sequence information.  A number of such tls have been publised in the
literature, operating on a number of different principles.  Tools such
as RECON and PILER are based on self-alignment, using WU-BLAST and
PALS for the alignemnt [CITE].  RECON show good sensitivity but is
computationally intesive and infeasible for use on whole genomes (requiring 60 hours for 18Mb rice genome in a 2013
report), while PILER achieves a good runtime with very low
sensitivey. [CITE JIANG].   ReAS and RepeatScout are based on $k$-mer
searches [CITE], with the earlier showing less sensitivity than
RECON.  Finally RepeatGluer is based on a vairation od DeBrujin
grauphs [CITE], which allows for a decomposition of TE families into
domains, but is very, very computationally expensive.  In Saha {\it et
  al.} the authors perform an extenive comparison of the tools, and
conclude that RepeatScout is the best tool overall for assembled
genomes and ReAS the best when dealing with unassembled sequeunce
fragments [CITE].

Another line of developement, first proposed by Zheng and Lonardi,
involves the use of {\it elementary repeats} [CITE].  Similar in concept to
the motifes identified by RepeatGluer [CITE], an elementary repeat is
foramlly defined sequence pattern that could be considered a buiding
block (or motif) of a transposable element.  Zheng and Lonardi defined
the concept and proposed an idenficiation algorith, but that algorithm
suffered from a runtime that was quadratic in the size of the genome.
Two other works, He [CITE] and Huo {\it et al.} employed variations of
suffix trees to get linear identification algorithms, but lost
sensitivity in the resence of genomic substitions -- making them less
useful on real data.

Here we describe RAIDER (Rapid Ab Initio Detection of Elementary
Repeats),

### 2015-07-22 manuscript/defs2.tex
Updating defs / manuscript

\newtheorem{definition}{Definition}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}

@@

\subsection{REFDs}
In RAIDER, everything is defined in terms of $l$-mers.  hHence we need
to generlize slightly to allow the ambiguity character, so we will
create a description of a {\it repeat element family discriptior},
with an $l$-mer being a special case of that.

@@

@@

For example: if we have $G = \ldots AAAAACTTTTT\ldots AAAAAGTTTTT
\ldots$ (notice: 6 As), with the seed $1101$, we get the refd $AAAAAA*TT*TT$, which
deposes into $AA*AA$, $AA*TT$, and $TT*T$$.  If any string compatible
with any of these three occurs anywhere else, it breaks the family.
But if the string $AA*AC$ occurs somewhere else, it does not -- thats
irrelevant.  (In this example the minimal decomposition is unique -- I'm not
sure that this is always the case, or what multiple decompositions
mean.)

@@

@@

I want to get a definition of elementary repeats that paralleles the
one we had.  (And, hopefully, is provably equivilent for spaced seeds
without spaces.)  This is tricky: I'm not yet sure what the equivilent
is for the minimality condition (there is no subsequence of an
elementary repeat that occurs independently of the instances of the
elementary repeat).  Lets look why.

Suppose we have in the string out genome $AAAAATCCCCC$ and
$AAAAAGCCCCC$.  Using the seed $s=11011$, we should get the refd
$r=AA*AA*CC*CC$. Here are different ways of phrasing the minimality
condition, and why they break:
\begin{enumerate}
\item No sub-refd can match any string in the genome outside of those
  matching $r$.  Doesn't work: we can have an AAAAT occuring somewhere
  else wihtout breaking this family, since the second $l$-mer is
  irrelevant to construction.
\item There cannot be any string $S'$, $|s'| < |r|$, occuring outside
  the instances, such that $s'$ generalizes to a sub-refd $s$ or $r$.
\end{enumerate}

\begin{definition}
Let $s$ be a fixed seed with length $l$, $f$ be a fixed integer, and $G$ be a fixed
genomic sequence.  An
\underline{Elementary Repeat Family} is a tuple $(r,S)$ where: 
\begin{enumerate}
\item $r$ is an refd such that $s$ covers $r$. 
\item $S$ is a set of genomic coordinates such that $|S| \geq f$.  Say
  that the sub-refds induced by $S$ are the substrings of $G$ starting
  at cordinates of $S$ and extended $|r|$ bases.
\item For every $i \in S$, $r$ is consistent with the string
  $G_{i:i+|r|}$.
\item Minimality condition.  Possibilities (I'm not sure which is
  right):
  \begin{itemize}
  \item There is no sub-refd $r'$ of $r$ that is covered by $s$ and is
     consistent some sequence $x$ that is sits outside the sequences
      induced by $S$.
  \item There is no $l$-mer $r'$ contained in $r$ that is consistence
    with $s$, such that $\sigma_s(r)'$ occurs anywhere in $G$ outside
    of the sequences induced by $S$.
 \end{itemize}
\item There is no refd $r'$ containing $r$ such that every sequence
  induced by $S$ is contained by some sequence consisted with $r'$.
\end{enumerate}
\end{definition}

{\bf Comments:} For the last two points, I'm really not sure if the
substring relation is the correct relation between $r$ and $r'$.

\subsection{Lemmas}

Right now, I'm trying trying to think ofpossibly interesting
statements and prove them.

Assume a fixed seed $s$ of length $l$ and a fixed genome $G$.

@@

>>>>

\newtheorem{definition}{Definition}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

@@

\subsection{REFDs}
In RAIDER, everything is defined in terms of $l$-mers.  Hence we need
to generalize slightly to allow the ambiguity character, so we will
create a description of a {\it repeat element family description},
with an $l$-mer being a special case of that.

@@

\begin{definition}
Given aan refd and a sequence $G$, let $freq_G(r)$ be the number of
positions $i$ at which $r$ generalizes $G_{i:i+|r|}$.
\end{definition}

{\bf Example:} Let $R=AA*AA$ and $G=AAAAACCAAGAAC$.  Then $freq_G(r)
= 2$.

@@

For example: if we have $G = \ldots AAAAACTTTTT\ldots AAAAAGTTTTT
\ldots$ (notice: 6 As), with the seed $1101$, we get the refd $AAAAAA*TT*TT$, which
deposes into $AA*AA$, $AA*TT$, and $TT*T$.  If any string compatible
with any of these three occurs anywhere else, it breaks the family.
But if the string $AA*AC$ occurs somewhere else, it does not -- thats
irrelevant.  (In this example the minimal decomposition is unique -- I'm not
sure that this is always the case, or what multiple decompositions
mean.)

@@

\begin{definition}
A decomposition $S$ of $r$ w.r.t. $s$ is \underline{maximal} it
countains every $i$ such that $s$ hits $r$ at $i$.
\end{definition}
(Clearly this is unique.)

Eample: 
\begin{itemize}
\item If $s=11011$ and $r=AAAAAA*CCCCCC$, then $D=\{0,1,4,7,8\}$.
\item If $s$ has no 0s, then $D = \{i \; : \l 0 \leq i \leq
  |r|-|s|\}$.
\end{itemize}

\begin{definition}
Given an refd $r$, a length $l$ covering seed $s$, and a decomposition $D$, the
\underline{sequence decomposion} $Q = \{\gamma(r_{i:i+l}) \; : \; i \in D}$ 
\end{definition}

{\bf Examples:}
\begin{itemize}
\item $s=11011$, $r=AAAAA*CCCCC$, the only minimal decomposition is
  $D=\{0,4,6\}$, and the corresponding sequence decomposition is
  $\{AA*AA, AA*CC, CC*CC\}$.
\item $s=11011$, $r=AAAAATCCCCC$, there are five different minimal
  decompositions.  For $\{0,1,6\}$ the sequence decomposition would be
  $AA*AA, AA*AT$, and $TT*TT$.
\end{itemize}

\begin{definition}
The {\it representitive sequence} of a decomposition is
created by compressing the overlapping portions of the elements of the
decompresion.
\end{definition}

{\bf Examples:}
\begin{itemize}
\item $s=11011$, $r=AAAAA*CCCCC$, $D={0, 3, 6}$,
  $Q=\{AA*AA,AA*CC,CC*CC$}.  The representative sequence would be
  $\{AA*AA*CC*CC\}$.
\item $s=11011$, $r=AAAAATCCCCC$, $D={0,2,6}$, $Q=\{AA*AA, AA*TC, CC*CC\}$.  
 The representative sequence would be $\{AAAAATCC*CC\}$.  We have to
 compress the overlap at $r=2:5$, which endsup removing both *
 symbols.
\end{itemize}

{\bf TO CONSIDER: What is the relationship between $r$ and the
  representative sequence of one of its decompostion?  Always a
  generlization of $r$?}

@@

\begin{definition}
For a ferquenct $f$, a genome $G$ and seed $s$, an refd $r$ is an \undetline{elementary
  repeat} if the following holds.  
\begin{enumerate}
\item $s$ covers $r$.
\item $freq_G(r) \geq f$.
\item For the maximimal sequence decomposition $Q$, 
  $freq_G(q) = freq_G(r)$ for all $q \in Q$.
\item There does not exist an $r'$ that is covered by $s$ such that
  (1) $r$ is compatable(2) with a sub-refd of $r'$, and (2) $r'$ meets
  conditions (1)-(3) of this definition.
\end{enumerate}
\end{definition}

\begin{theorm}
If $s$ is a contiguous spaced seed (that is, no 0s) and $r$ is a
repeat element w.r.t to $s$, the $r$ is a repeat element by Definition
5 of Nate's thesis.
\end{theorem}
{\bf Proof:} Let $s$ be a spaces seed of length $L$ with no 0 symbols,
and $r$ be an elementary repeat.  Note that $r$ has no $*$
symbols, as otherwise $s$ could not cover it.  Now consider the
conditions for Definition 5:
\begin{itemize}
\item The repeat element is at least length $L$: since $r$ is covered
  by seed $s$ of length $L$, this is clearly true.\footnote{There
    appears to be a mistake in Nate's thesis with condition 1 -- this
    is what it was supposed to mean.}
\item $freq_G(r) \geq f$: Follows directly from (2).
\item Every substring of $r$ has the same frequence as $r$: Since $s$
  has no 0s, $Q$ is the set of all substrings of $r$. Hence this
  condition applies to every substring of $r$, which is what we nee.
\item $r$ is maximal w.r.t to Nate's defintion: Suppose this were not
  true: we could add 1 character to $r$  to form $r'$, and have an element that
  meats conditions Nate's definition.  (Note that, with no spaces, if
  we can add one or more characters we can add only 1 character -- we
  can assume are only adding 1.)  That means that every subsequence of
  $r'$ has the same frequency in $G$ (by condition (3) of Nate's
  defintion), meaning that they fufill condition (3) of our
  definition.  The other two conditions of our definition our
  immediate, meaning $r'$ contradicts (4) and hence $r$ is not an elementary
  repeat (by our definition).
\end{itemize}

\subsection{Lemmas}
Assume a fixed seed $s$ of length $l$ and a fixed genome $G$.

\begin{definition}
An $l$-mer {\it belongs} to a repeat element $r$ if the element is
compatable with an element in the maximal sequence decomposition of $r$.

\begin{lemma}
An $l$-mer can belong to at most one elementary repeat family.
\end{lemma}
Using our definition of ``belongs'', violating this would vilate
condition (3) of the r.e. definition.

\begin{observation}
For any string $l$-mer $w$, if $freq(\gamma_s(w)) \geq w$ than
$\gamma_s(w)$ is either is an elementary
repeat or is compatable with some substring of an elementary repeat.
\end{observation}
$\gamma_s(w)$ trivially satsfies the first three conditions, so only
the fourth condition could break it.

\begin{lemma}
Let $s_1$ and $s_2$ be two sequences in $G$ such that refd generalizes
both but there is no refd that generalizes generlizes superstrings
properly containing each, then $r$ is not compatable with a sub-refd
of any repeat element family.
\end{lemma}
Violation of this would imply a violation of (2) in the definition.

\begin{lemma}
If two $l$-mers belong to the same elementary repeat, then they have
the same frequency.
\end{lemma}
Given our definition of ``belongs'', this follow immediately.

@@

%%  LocalWords:  REFDs mers

### 2015-07-22 manuscript/defs3.tex
Updating defs / manuscript

\subsection{REFDs}

@@

We will allow any normal string operator to apply to sequence descriptors.  In specific, for a s.d. $r$:
\begin{itemize}
\item $r_{i,i+l}$, or $r[i:i+l]$, is the length $l$ sub-sd (substring) of $r$ starting at $i$.
\item $|r|$ is the number of characters in $r$.
\item $||r||$ is the number of non-* characters in $r$.
\item We will use $l$-mer to denote a length $l$ sd.
\end{itemize}

\begin{definition}
  If $r_1$ and $r_2$ are sds of the same length:
\begin{enumerate}
\item $r_1$ and $r_2$ are \und{consistent} if, for any position $i$ in
  which they differ, one of them is a $*$. 
\item $r_2$ \und{generalizes} $r_1$ (and $r_1$ is \und{generlized} by $r_2$),
  if, for any position in which they are different, $r_2$ has a $*$.
  \end{enumerate}
\end{definition}

\begin{definition}
  The frequency of an s $r$ in genome sequence $G$ is the number of
  $i$ such that $G_{i:i+|r|}$ is compatible with $r$.  
\end{definition}

@@

\subsection{REFDs and Spaced Seeds}

@@

  \begin{itemize}
  \item We say $s$ and an $l$-mer $r$ are \und{consistent} if, when $s_i = 0$, $r_i = *$.
  \item We say seed $s$ \und{hits} sd $r$ at position $i$ such that
    $s$ is consistent with $r_{i:i+|s|}$ ($i \leq |r| - |s|$).
  \item We say seed $s$ \und{covers} sd $r$ if, for every $j$, $0 \leq
    j < |r|$, there is an $i$ such that $s$ hits $r$ at $i$ and covers
    position $j$ in doing so.
  \end{itemize}
\end{definition}

Example: For $s=11011$ and $r=AAAAA*TTTTT$, $s$ covers $r$, hitting at
$\{0,3,6\}$.  $s$ does not cover $AAAAA**TTTTT$, as there is no
position at which $s$ hits and extens over $i=5$.

@@

\begin{notation}
  \begin{itemize}
  \item Let $\gamma_s(r)$ denote the generalization of $r$ w.r.t. $s$
  (undefined if $|r| \neq |s|$).
  \item Let $\delta_s(r)$ denote the decomposition of $r$ w.r.t. $s$.
  \item Let $\Delta_s(r)$ denote the sequence decomposition of $r$ w.r.t. to
  $s$.
  \item Let $\Gamma_s(r)$ denote the generalized seqeunce decomposition of
  $r$ w.r.t. $s$.
  \end{itemize}
\end{notation}

@@

\subsection{Elementary repeats}

@@

\begin{definition}
  A {\it repeat element family} is a tuple $(r,S)$ where $r$ is a
  repeat element descriptor and $S$ is the set of coordinates of $G$
  where there is a sequence compatiable with $r$.
\end{definition}

\subsection{Equivilance proof}

@@

\begin{definition}
\label{NateDef}
Given a genomic sequence $G$ and integers $f$ and $l$, a seqeunce $r$
is an {\it elementary repeat} if we can write $r = x_0 \cdot x_2 \cdot
\cdot x_k$ $where $|x_i|=l$, $k = |r|-|l|$, and:
\begin{enumerate}
\item $k \geq 0$.  (There is at least one $l$-mer substring).
\item $freq_G(r) \geq f$
\item $freq_G(x_i) = freq_G(r)$
\item There does not exist an $y$, $|y|=l$ such that either $y \cdot
  r$ or $r \cdot y$ meets conditions 1-3.
\end{enuemrate}
\end{definition}
Recall that $x \circ y$ is defined only when $x$ has an $|x|-1$ length suffix
that is a prefix of $y$, and represents their ``merging''.
(e.g. $ACCC \circ CCCG = ACCCG$.)

\begin{theorem}
  Consider a genome $G$, an integer $f$, and a seed $s$ such that $s$
  has no zero chartacters.  $r$ is an elementary
  repeat descriptor for $s$ (by Definition~\label{ERdef}) if and only if $r$
  is an elementary repeat for $l=|s|$ (by
  Definition~\label{NateDef})).
\end{theorem}

@@

($Leftarrow$) First we observe that if $s$ has no zeros, and covers
$r$, then $r$ has no * characters.  From this it follows that
$\Gamma_s(r) = \{r_0,r_1,\cdtos,r_{k}$ where $|r_i|=l$ and $k=r-l$ --
every $l$-mer contained with $r$.  From this it follows that $r = r_0 \circ
r_1 \

Now we prove each of the four conditions of Nate's definiciton:
\begin{enumerate}
  \item Since $r$ is covered by $s$, $|r| \geq |s|$, so $k =
    |r|-|s|+1 > 0$.
  \item Since $r$ contains no * symbols, the only string it
    is compatible with is itself.  Meaning it has the sme frequency by
    either definition (sd or string).  Hence (2) of Nate's
    definition following directly from (2) of out definition.
  \item Follows immediately from the fact that every $r_i$is a
    member of $\Delta_G(r)$, and hence must have the right frequency.
   \item Argue by contradiction: suppose there were such a $y$ (assume
     w.l.g. it follow $r$, and let $r' = r \circ x$.  Since Nate's
     definition tells us $r'$ matches a string around every occurance
     of a match to $r$, this means $r'$ is (trivially) compatable to a
     substring around every occurance of a match to $r'$ -- to $r'$
     could not have been a repeat element descriptor.
\end{enumerate}

($Rightarrow$) This is the same argument backwards, and not worth
writing out.
\end{enumerate}

\subsection{Basic Lemmas}

@@

{\bf COMMENT:} HMM...I'm worried about this one.  What if we extneded
by a whole sead length?  Could that matter?

\begin{lemma}
  If two $l$-mers appear blog to the same elementary repeat
  descriptor, then they have the same frequence.

@@

\subsection{Variations}

@@

For $11011$ and $G = \cdots ACCCCCGTTTTTA\cdots ACCCCCATTTTTA$, we can
have $AC*CCC*TT*TTA$ or $ACC*CC*TTT*TA$$, among others.

What do we know about the r.e.d. that Carly's algorithm will pick?

>>>>

\section{REFDs}

@@

We will allow any normal string operator to apply to sequence descriptors.  In specific, for an s.d. $r$:
\begin{itemize}
\item $r_{i,i+l}$, or $r[i:i+l]$, is the length $l$ sub-sd (substring) of $r$ starting at $i$.  (Python-like notation.)
\item $|r|$ is the number of characters in $r$.
\item $||r||$ is the number of non-* characters in $r$.
\item We will use $l$-mer to denote a length $l$ sd.
\end{itemize}

\begin{definition}
  If $r_1$ and $r_2$ are sds of the same length:
\begin{enumerate}
\item $r_1$ and $r_2$ are \und{consistent} if, for any position $i$ in
  which they differ, one of them has a $*$ in that position.
\item $r_2$ \und{generalizes} $r_1$ (and $r_1$ is \und{generlized} by $r_2$),
  if, for any position in which they are different, $r_2$ has a $*$.
  \end{enumerate}
\end{definition}

\begin{definition}
  The frequency of an sd $r$ in genome sequence $G$ is the number of
  $i$ such that $G_{i:i+|r|}$ is compatible with $r$.  
\end{definition}
{\bf Note:} Debated whether this should be ``compatible'' or ``generalized'' -- then realized that since $G$ is a string (as opposed to an sd) they are equivilent.

@@

\section{REFDs and Spaced Seeds}

@@

  \begin{itemize}
  \item We say $s$ and an $l$-mer $r$ are \und{consistent} if, when $r_i = *$, $s_i=0$.
  \item We say seed $s$ \und{hits} sd $r$ at position $i$ such that
    $s$ is consistent with $r_{i:i+|s|}$ ($i \leq |r| - |s|$).
  \item We say seed $s$ \und{covers} sd $r$ if, for every $j$, $0 \leq
    j < |r|$, there is an $i$ such that $s$ hits $r$ at $i$ and covers
    position $j$ in doing so.
  \end{itemize}
\end{definition}

Example: For $s=11011$ and $r=AAAAA*TTTTT$, $s$ covers $r$, hitting at
$\{0,3,6\}$.  $s$ does not cover $AAAAA**TTTTT$, as there is no
position at which $s$ hits and extends over $i=5$.

@@

\begin{notation}
Notation for the concepts from this section:
  \begin{itemize}
  \item Let $\gamma_s(r)$ denote the generalization of $r$ w.r.t. $s$
  (undefined if $|r| \neq |s|$).
  \item Let $\delta_s(r)$ denote the decomposition of $r$ w.r.t. $s$.
  \item Let $\Delta_s(r)$ denote the sequence decomposition of $r$ w.r.t. to
  $s$.
  \item Let $\Gamma_s(r)$ denote the generalized seqeunce decomposition of
  $r$ w.r.t. $s$.
  \end{itemize}
\end{notation}

@@

\section{Elementary repeats}

@@

\begin{definition}
  A {\it repeat element family} is a set $S$ of genome coordinates
  such that there exists an sd $r$ where:
\begin{itemize}
\item $r$ generlizes the string $G_{i:i+|r|}$ for every $i \inS4.
\item $r$ does not generalize any other substring of $G$.
\end{definition}
Notice that there can be multiple such $r$ for a given set.  Any two
such $r$ for the same $S$ will be compatible, but they don't
necessarily generlize each other.

\section{Equivilance proof}

@@

\begin{definition}
\label{NateDef}
Given a genomic sequence $G$ and integers $f$ and $l$, a seqeunce $r$
is an {\it elementary repeat} if we can write $r = x_0 \circ x_2 \circ
\cdots \circ x_k$ where $|x_i|=l$, $k = |r|-|l|$, and:
\begin{enumerate}
\item $k \geq 0$.  (There is at least one $l$-mer substring).
\item $freq_G(r) \geq f$
\item $freq_G(x_i) = freq_G(r)$
\item There does not exist an $y$, $|y|=l$ such that either $y \circ
  r$ or $r \circ y$ meets conditions 1-3.
\end{enumerate}
\end{definition}
Recall that $x \circ y$ is defined only when $x$ has an $|x|-1$ length suffix
that is a prefix of $y$, and represents their ``merging''.
(e.g. $ACCC \circ CCCG = ACCCG$.)

\begin{theorem}
  Consider a genome $G$, an integer $f$, and a seed $s$ such that $s$
  has no zero chartacters.  $r$ is an elementary
  repeat descriptor for $s$ (by Definition~\ref{ERdef}) if and only if $r$
  is an elementary repeat for $l=|s|$ (by
  Definition~\ref{NateDef})).
\end{theorem}

@@

($\Leftarrow$) First we observe that if $s$ has no zeros, and covers
$r$, then $r$ has no * characters.  From this it follows that
$\Gamma_s(r) = \{r_0,r_1,\cdots,r_{k}\}$ where $|r_i|=l$ and $k=r-l$.
That is, every $l$-mer contained with $r$.  From this it follows that $r = r_0 \circ
r_1 \circ \cdots \circ r_k$.

Now we prove each of the four conditions of Nate's definiciton:
\begin{enumerate}
  \item Since $r$ is covered by $s$, $|r| \geq |s|$, so $k =
    |r|-|s|+1 > 0$.
  \item Since $r$ contains no * symbols, the only string it
    is compatible with is itself.  Meaning it has the same frequency by
    either definition (sd or string).  Hence (2) of Nate's
    definition follows directly from (2) of our definition.
  \item Follows immediately from the fact that every $r_i$is a
    member of $\Delta_G(r)$, and hence must have the right frequency.
   \item Argue by contradiction: suppose there were such a $y$ (assume
     w.l.g. it follow $r$), and let $r' = r \circ x$.  Since Nate's
     definition tells us $r'$ matches a string around every occurance
     of a match to $r$, this means $r'$ is (trivially) compatable to a
     substring around every occurance of a match to $r'$ -- so $r'$
     could not have been a repeat element descriptor.
\end{enumerate}

($\Rightarrow$) This is the same argument backwards, and not worth
writing out.

\section{Basic Lemmas}

@@

{\bf COMMENT:} HMM...I'm worried about this one.  What if we extneded
by a whole se1d length?  Could that matter?

\begin{lemma}
  If two $l$-mers belong to the same elementary repeat
  descriptor, then they have the same frequency.

@@

\section{Variations}

@@

For $11011$ and $G = \cdots ACCCCCGTTTTTA\cdots ACCCCCATTTTTA$, we can
have $AC*CCC*TT*TTA$ or $ACC*CC*TTT*TA$, among others.

\subsection{Carly's output}

\begin{definition}
Given a repeat element family $S$ and a descriptor $r$, we say:
\begin{itemize}
\item We say position $i$ is variable in $S$ if there are two elements
  of $S$ whose sequences vary at positin $i$.
\item Position $i$ of $r$ is \und{tight} if either $r_i \neq *$ or
  $S$ variable at position $i$.  (That is, $r$ must have a *.)
\item We say $r$ is tight if $r_i$ is tight for all $0 \leq i < |r|$.
\end{itemize}
\end{definition}

Facts about Carly's output:
\begin{itemize}
\item If $S$ is invarient in all positions, then the results $r$ has
  no * characters.  For every $i$, $s$ hits at either $i$ or
  $i-|s|+1$, meaning that there is a $w \in \Gamme_s(r)$ that starts
  (or ends) at position $i$.  Since a seed starts (ends) with a 1,
  there canot be a * at this position.
\item The output descriptor is not necessarily tight.  For example:
  $s=11011$, where $G$ contains $AAAAACGGGGG$ and $AAAAATGGGGG$.  Then
  $r = AA*AA*AA*AA$, hence is not tight at position 2.
\end{itemize}
